Practical Optional Types for Clojure

### 2015-02-27 formal-model.tex
overview
tweak figure
some bib

Now that we have demonstrated the core features Typed Clojure
provides, we link them together in a formal model called
\lambdatc{}.
Our presentation will start with a review of
occurrence typing~\cite{TF10}.
Then for the rest of the section we incrementally add each
novel feature of Typed Clojure to the formalism,
interleaving presentation of typing rules, operational semantics
and subtyping.

The first insight about occurrence typing is that
propositions can be used to capture type information about our programs,
in particular by referencing components or subcomponents of the runtime environment
and relating them to types.
\emph{Type Propositions} \prop{} make assertions like ``variable \clj{x} is of type \clj{Number}'' or
``variable \clj{x} is not \clj{nil}''---
in our logical system we write these as
{\isprop{\NumberFull}{\x{}}}
and {\notprop{\Nil{}}{\x{}}}. 
The other propositions are standard: the implication, conjunction,
disjunction, trivial (\topprop{}) and impossible (\botprop{}) propositions
(figure~\ref{main:figure:termsyntax}).

@@

\subsection{Paths}

@@

    \tatypesyntax{}\\
    \typehintenvsyntax{}\\
    \classtableallsyntax{}
  \end{altgrammar}
  $$
 \classtablelookupsyntax{}
  \convertjavatypegrammar{}
  \begin{mathpar}
    {\TNewStatic}
    {\TFieldStatic}

    {\TMethodStatic}
  \end{mathpar}
  \begin{mathpar}
    \BField{}

>>>>

Now that we have demonstrated the core features Typed Clojure
provides, we link them together in a formal model called
\lambdatc{}.
Our presentation will start with a review of
occurrence typing~\cite{TF10}.
Then for the rest of the section we incrementally add each
novel feature of Typed Clojure to the formalism,
interleaving presentation of syntax, typing rules, operational semantics
and subtyping.

%TODO EXPLAIN what is a proposition?
The first insight about occurrence typing is that
logical formulas
can be used to represent type information about our programs
by relating parts of the runtime environment to types.
Occurrence typing uses propositional logic 
\emph{Type Propositions} \prop{} make assertions like ``variable \clj{x} is of type \clj{Number}'' or
``variable \clj{x} is not \clj{nil}''---
in our logical system we write these as
{\isprop{\NumberFull}{\x{}}}
and {\notprop{\Nil{}}{\x{}}}. 
The other propositions are standard logical connectives: implications, conjunctions,
disjunctions, and the trivial (\topprop{}) and impossible (\botprop{}) propositions
(figure~\ref{main:figure:termsyntax}).

@@

\subsection{Paths}
\label{sec:formalpaths}

@@

    \tatypesyntax{}\\
    \typehintenvsyntax{}\\
    \classtableallsyntax{}
  \end{altgrammar}
  $$
 \classtablelookupsyntax{}
 \begin{mathpar}
  \begin{altgrammar}
    \convertjavatypenil{}
  \end{altgrammar}
  \begin{altgrammar}
    \convertjavatypenonnil{}
  \end{altgrammar}
\end{mathpar}
  \begin{mathpar}
    {\TNewStatic}
    {\TFieldStatic}

    {\TMethodStatic}

    %{\TInstance}
  \end{mathpar}
  \begin{mathpar}
    \BField{}

### 2015-02-27 intro.tex
overview, Thursday evening changes
some bib

\begin{figure}
\inputminted[firstline=5,lastline=8]{clojure}{code/demo/src/demo/parent2.clj}
\caption{A simple Typed Clojure program}
\label{fig:ex1}
\end{figure}

Figure~\ref{fig:ex1} presents a simple program demonstrating many
aspects of our system, from simple type annotations to explicit
handling of Java's \java{null} in interoperation, as well as an
extended form of occurrence typing and Clojure's \emph{type hints},
which are central to Typed Clojure's approach to interoperability.

>>>>

\begin{figure}
\inputminted[firstline=5,lastline=7]{clojure}{code/demo/src/demo/parent2.clj}
\caption{A simple Typed Clojure program}
\label{fig:ex1}
\end{figure}

Figure~\ref{fig:ex1} presents a simple program demonstrating many
aspects of our system, from simple type annotations to explicit
handling of Java's \java{null} (written \clj{nil}) in interoperation, as well as an
extended form of occurrence typing and Clojure's \emph{type hints},
which are central to Typed Clojure's approach to interoperability.
